Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-08-15 document/ms.tex
some more text to the addressing binarity subsection in discussion

The IGRINS spectrum demonstrates some features that are reproduced only in the hot photosphere model, or only in the cool photosphere model.  Figure \todo{FigureRef} shows a selection of such spectral features for a range of fill factors.  The temperature-sensitivity for the whole-order spectral fit is shown as a function of fill factor in the rightmost panels.  The unshown portion of the spectral order dilutes the temperature-sensitivity of these distinctive features, but overall the fill factor correlates with effective temperature in order to reproduce the observed line-depth ratio.  In some cases, featureless cool photosphere spectra veil isolated spectral lines predicted in the hot photosphere models.  In other cases, the hot photosphere model veils sequences of shallow spectral features predicted in the cool photosphere model.  The cool photosphere models predict shallow features because line blanketing from multiple indistinct molecular bands overlap from rotationally broadening.  This line blanketing has probably hampered efforts to identify isolated spectral features suitable for line-depth-ratio analysis in the spectra of rotationally broadened young stars.  One such feature has been previously identified in the near-IR: the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in our data, with the central line exceeding the depth of the adjacent two lines, although the \PHOENIX\ models predict a non-negligible hot star contribution to the middle line at our range of effective temperatures.  In comparison, the forward-modeling technique described in this work thrives in the presence of long sequences of indistinct yet predictably correlated spectral features.  Similar forward-modelling strategies have successfully identified weak metal lines in the spectra of extremely low metallicity stars \todo{cite-Evan Kirby}.  Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.

@@

 \includegraphics[width=0.80\textwidth]{figures/spectral_postage_stamp_03} 
 \caption{Examples of spectral features in the observed IGRINS spectrum (thick gray line) that are present only in the hot (blue line) photosphere model (top panel), only in the cool (red line) photosphere model (middle panel), or in a mixture of hot and cool models (bottom panel).  The composite spectrum model (purple thin line) is consistent with the observed spectrum for a range of fill factors, with examples of the the median fill factor (middle panel spectrum) and $\pm2\sigma$ fill factors demarcated on the spectral postage stamps.  The right-most panels show the temperature-fill factor correlation for these spectral orders, with a straight-line fit (dashed black line) through all the samples.  The location of the samples used to make the synthetic spectra to the left are shown as solid black symbols.}
 \label{fig:specPostageStamps}
\end{figure*}

@@

\emph{Suitability of pre-computed \PHOENIX\ model spectra, model ranges, $\A \neq 0$, and magnetic fields-} We have assumed that the \PHOENIX\ synthetic spectra are a good representation of the observed IGRINS spectrum of \name, a young star likely to possess a large magnetic field.  Cool sunspots are associated with heightened magnetic fields on the sun.  Magnetic fields will have two effects on the emergent spectrum- line broadening attributable to the Zeeman effect, and pressure broadening attributable to seeing deeper into deeper layers in the photosphere compared to the surrounding hotter photosphere.  The former effect will differentially impact lines with high magnetic sensitivity \citep{2013AJ....146...51D}.  The latter effect could masquerade as an apparent shift in surface gravity, as seen in mixture models of spotted stars with empirical templates composed of dwarfs and giants \citep{oneal96}.  Our tacit assumptions have been that these effects are 1) relatively small compared to the gross appearance and disappearance of spectral lines at disparate temperatures, 2) there exist many orders and lines that are not affected at all, and 3) the Gaussian process noise model allows some resilience against unknown residual spectrum correlations.  We see tentative evidence for physics outside of the pre-computed model grids in the poor fits to the observed spectra in $K-$band.  Strong metal lines demonstrate large departures in their best fit $\Z$ due to Zeeman broadening.  No pre-computed model grids including a range of surface magnetic fields are available at the wavelength range and spectral sampling needed to approach these questions within this spectral inference framework.

@@

\subsection{Evidence for intra-spectrum RV jitter}
Starspots cause significant jitter in radial velocity surveys for young planets \citep[e.g.][]{donati14, robertson14}.  The two-temperature model assumed that the hot and cool components shared the same $\vsini$ and $v_z$.  This assumption is accurate for either homogeneously distributed spots, azimuthally symmetric bands of spots, or polar spots.  However, LkCa 4 has clear optical line-profile variations \citep{nguyen12, donati14}.  Interestingly, we should be able to detect anticorrelation between lines attributable to hot photosphere, and lines attributable to cool photosphere, since there is a zero-sum competition for solid-angle on the recessional and advancing stellar limbs.  We found tentative, albeit weak, evidence for such anticorrelation in $v_z$ across spectral orders with dominated by cool photosphere and those dominated by hot photosphere.  For example, the cool-photosphere dominstated spectral orders $o = 100, 102, and 104$ show the largest median $v_z\sim14.5-17.0$ km/s, while the other orders that possess a mix of primarily hot- and some cool- photosphere features demonstrate $v_z\sim9-13$ km/s.

Radial velocity (RV) planet searches in the infrared--where the modulation of RV signals is lessened \citep{prato08}-- could potentially reduce RV jitter by characterizing and constraining the effect of direct emission from starspots.

@@

\subsection{Addressing binarity}
In principle, unresolved eclipsing binaries could mimic some of the data shown here.  However, unresolved such a system would have had to evade detection in high-resolution imaging \citep{kraus11,daemgen15}...

>>>>

The IGRINS spectrum demonstrates some features that are reproduced only in the hot photosphere model, or only in the cool photosphere model.  Figure \ref{fig:specPostageStamp} shows a selection of such spectral features for a range of fill factors.  The temperature-sensitivity for the whole-order spectral fit is shown as a function of fill factor in the rightmost panels.  The unshown portion of the spectral order dilutes the temperature-sensitivity of these distinctive features, but overall the fill factor correlates with effective temperature in order to reproduce the observed line-depth ratio.  In some cases, featureless cool photosphere spectra veil isolated spectral lines predicted in the hot photosphere models.  In other cases, the hot photosphere model veils sequences of shallow spectral features predicted in the cool photosphere model.  The cool photosphere models predict shallow features because line blanketing from multiple indistinct molecular bands overlap from rotationally broadening.  This line blanketing has probably hampered efforts to identify isolated spectral features suitable for line-depth-ratio analysis in the spectra of rotationally broadened young stars.  One such feature has been previously identified in the near-IR: the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in our data, with the central line exceeding the depth of the adjacent two lines, although the \PHOENIX\ models predict a non-negligible hot star contribution to the middle line at our range of effective temperatures.  In comparison, the forward-modeling technique described in this work thrives in the presence of long sequences of indistinct yet predictably correlated spectral features.  Similar forward-modelling strategies have successfully identified weak metal lines in the spectra of extremely low metallicity stars \todo{cite-Evan Kirby}.  Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.

@@

 \includegraphics[width=0.80\textwidth]{figures/spectral_postage_stamp_03} 
 \caption{Examples of spectral features in the observed IGRINS spectrum (thick gray line) that are present only in the hot (blue line) photosphere model (top panel), only in the cool (red line) photosphere model (middle panel), or in a mixture of hot and cool models (bottom panel).  The composite spectrum model (purple thin line) is consistent with the observed spectrum for a range of fill factors, with examples of the the median fill factor (middle panel spectrum) and $\pm2\sigma$ fill factors demarcated on the spectral postage stamps.  The right-most panels show the temperature-fill factor correlation for these spectral orders, with a straight-line fit (dashed black line) through all the samples.  The location of the samples used to make the synthetic spectra to the left are shown as solid black symbols.}
 \label{fig:specPostageStamp}
\end{figure*}

@@

We assumed that the \PHOENIX\ synthetic spectra are a good representation of the observed IGRINS spectrum of \name, a young star likely to possess a large magnetic field.  Cool sunspots are associated with heightened magnetic fields on the sun.  Magnetic fields will have two effects on the emergent spectrum- line broadening attributable to the Zeeman effect, and pressure broadening attributable to seeing deeper into deeper layers in the photosphere compared to the surrounding hotter photosphere.  The former effect will differentially impact lines with high magnetic sensitivity \citep{2013AJ....146...51D}.  The latter effect could masquerade as an apparent shift in surface gravity, as seen in mixture models of spotted stars with empirical templates composed of dwarfs and giants \citep{oneal96}.  Our tacit assumptions have been that these effects are 1) relatively small compared to the gross appearance and disappearance of spectral lines at disparate temperatures, 2) there exist many orders and lines that are not affected at all, and 3) the Gaussian process noise model fosters some resilience against unknown residual spectrum correlations.  We see tentative evidence for physics outside of the pre-computed model grids in the poor fits to the observed spectra in $K-$band.  Strong metal lines demonstrate large departures in their best fit $\Z$ due to Zeeman broadening.  No pre-computed model grids including a range of surface magnetic fields are available at the wavelength range and spectral sampling needed to approach these questions within this spectral inference framework.

Uncertainties in the \PHOENIX\ synthetic spectra continuum opacity flow down to uncertainties in the absolute value of the fill factor.  The reliability of the stellar model continuum opacities has been called into question (A. Mann, \emph{priv. comm.}).  Erroneous bulk opacities could be responsible for some of the scatter of fill factor distributions seen across different spectral orders, as in Figures \ref{fig:TwoTempResults} and \ref{fig:specPostageStamp}.

@@

\subsection{Evidence for intra-spectrum RV jitter}
Starspots cause significant jitter in radial velocity surveys for young planets \citep[e.g.][]{donati14, robertson14}.  The two-temperature model assumed that the hot and cool components shared the same $\vsini$ and $v_z$.  This assumption is accurate for either homogeneously distributed spots, azimuthally symmetric bands of spots, or polar spots.  LkCa 4 has clear optical line-profile variations \citep{nguyen12, donati14}.  Accordingly, we should be able to detect anticorrelation between lines attributable to hot photosphere, and lines attributable to cool photosphere, since there is a zero-sum competition for solid-angle on the recessional and advancing stellar limbs.  We found tentative, albeit weak, evidence for such anticorrelation in $v_z$ across spectral orders dominated by cool photosphere and those dominated by hot photosphere.  For example, the cool-photosphere dominated spectral orders $o = 100, 102, and 104$ show the largest median $v_z\sim14.5-17.0$ km/s, while the other orders that possess a mix of primarily hot- and some cool- photosphere features demonstrate $v_z\sim9-13$ km/s.

Radial velocity (RV) planet searches in the infrared--where the modulation of RV signals is lessened \citep{prato08}-- could potentially reduce RV jitter by characterizing and constraining the effect of direct emission from starspots, in addition to the flux deficits imbued in optical line profiles.  Direct spectral emission from cool starspots is likely much weaker in sources with less starspot coverage than LkCa 4.

@@

\subsection{Addressing binarity}
In principle, unresolved binarity could mimic aspects of the data shown here, like the near-IR SED.  However, such a system would have had to evade detection in high-resolution imaging \citep{kraus11,daemgen15} and spectroscopic monitoring \citep{nguyen12,donati14}.  The sources would need to be eclipsing in order to yield $V-$band variability, which does not indicate a regular eclipsing binary signal in the light-curve (\emph{c.f.} Figure \ref{fig:PhotPhase}).  These hypothetical binaries would also need the same rotational broadening, since the IGRINS spectrum shows comparable estimates for $\vsini$ for both the cool and hot components.  So all of these lines of evidence render a hypothetical close eclipsing binary unviable.
